\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgments}

This dissertation is the result of my research at the Institute of Information Management at ETH Zurich, the Bosch Internet of Things Services Lab, the Human Dynamics Group and the Digital Currency Initiative at the MIT Media Lab. I would like to express my sincere gratitude to all those who so greatly supported me in these projects and made this work possible.
In particular I would like to thank my supervisor Prof. Dr. Elgar Fleisch for giving me the opportunity to benefit from the unique and stimulating research environment he has created. Furthermore, I'm very thankful for his constant support to explore my ideas. I would like to thank Prof.\ Dr.\ Fr\'{e}d\'{e}ric Thiesse for his willingness to co-supervise my research, his interest in my work and the valuable feedback he provided. I would like to thank Ass.-Prof. Dr. Felix Wortmann for his guidance throughout my research, and I am grateful for the many things I learned from him. I would like to thank Prof. Dr. Markus Weinberger and Timo Gessmann for supporting me and my projects from the industry perspective. I would like to thank Prof. Alex \emph{Sandy} Pentland for inviting me to spend 8 months at his research group at the MIT Media Lab. I also would like to thank the Digital Currency Initiative, in particular Joi Ito, Dr. Neha Narula, and Michael Casey for a stimulating environment and a welcoming and supportive atmosphere. 

I would like to thank the Bosch Group and Bosch Software Innovations in particular, whose commitment to research in the Internet of Things provided the foundation of this work. Furthermore, I would like to thank the Swiss National Science Foundation for supporting my research stay at the MIT Media Lab.

I would like to thank my colleagues at and around the Institute of Information Management at ETH Zurich and the Institute of Technology Management at the University of St. Gallen for the excellent working atmosphere and the many memorable hours we spent together. In particular, I would like to thank Thomas von Bomhard, Paul Rigger, Kristina Fl{\"u}chter, Marcus K{\"o}hler, Stefanie Turber, Andr{\'e} Dahlinger, Benjamin Ryder, Bernhard Gahr, Matthieu Chanson, Andreas Bogner, Arne Meeuw, Dirk Volland and Kay Noyen. Furthermore, I would like to thank Christian Decker, Francisc Bungiu, and Andrew Koh for insightful collaboration.



\endgroup
